An Extensible Approach to Verification of Embedded Network Systems*

### 2014-04-21 tech-project/paper.tex
adding information on our new shiny external syntax. I love the smell of new syntax in the morning

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of TinyOS applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of a toggling capability enhancement that we demonstrate running on a Telos mote.
\end{abstract}

@@

\subsection{Concepts}

A \textit{concept} module in RESOLVE defines a specification for a mathematical, abstract type. Similar to an interface in Java, a concept provides a number of operation signatures that implementors are expected to realize. Shown below is a \texttt{LED\_Template} concept that provides a light strip abstraction.

@@

It's worth noting that unlike the \texttt{LED\_Template} presented in \cite{regula:2010} which models an LED as the \textit{cartesian product} of booleans $b_1$, $b_2$, \ldots , $b_4$, the strip model we present here instead uses strings for the following reasons:
\begin{itemize}
\item Strings are indexible, and thus do not require separate \texttt{Set} and \texttt{Status} operations for each individual LED.
\item This approach demonstrates the benefits of reusable mathematical theories. The specifications here are based solely in \texttt{String\_Theory}, a pre-existing RESOLVE math library that provides all the definitions and theorems used in this specification.
\end{itemize}

The concept provides two operations. The first operation, \texttt{Set}, takes as a parameter an instance of an LED strip \texttt{L}, a boolean \texttt{b}, and an integer \texttt{i}. The operation \texttt{requires} that \texttt{i} falls within the length of the strip, and \texttt{ensures} upon completion of the operation that the length of the outgoing strip \texttt{L} is the same length as the incoming strip, \texttt{\#L}, and that the LED in position \texttt{i} of \texttt{L} is set to boolean \texttt{b}. The \texttt{Status} operation is specified similarly.

@@

%\subsection{An External LED Realization}

%Armed with a specification, the RESOLVE programmer would now ideally provide a verifiable implementation of \texttt{LEDs\_Template} via a RESOLVE realization. However, as our target platform is embedded, and our concept  aims to provide control for LEDs -- a decidedly low level feature -- our realization is forced to operate at similarly low levels by manually interfacing with hardware pins provided by msp430 chipset. RESOLVE, however in its current state is far too high level of a language to perform these tasks directly -- that is, it currently lacks driver support to do so. To address this difficulty, we introduce language support for \textit{external realizations}, which, in essence, leaves the user free to write their own realization of a concept in whatever language is most convenient. In the case of \texttt{LEDs\_Template}, we provide a hand written \texttt{Telos\_LEDs\_Realiz} C realization that matches as closely as possible the model of translation detailed in Section \#.

@@

Even enhancement realizations are neutral since only the operations called are those provided by the base concept -- and thus require no knowledge of how the called operation is implemented.

@@

\begin{figure}[!htb]
\centering
\begin{tabular}{lccc}
	\toprule
	Condition \# & Time (ms)	& Steps	& Search \\
	\midrule
	VC 0\_1	& 4426	& 5	& 0	\\
	VC 0\_2	& 5039	& 5	& 0	\\
	VC 0\_3	& 6324	& 6	& 0	\\
	\bottomrule
\end{tabular}
\caption{Verification results for \texttt{Toggle} (left)}
\label{fig:results}
\end{figure}

The results of verification are summarized Figure \ref{fig:results}. Using the RESOLVE compiler's integrated prover, we are able to mechanically and automatically dispatch all VCs generated by operation \texttt{Toggle}, thus verifying its correctness. In terms of proof difficulty, given the number of steps and time taken to establish each, leads us to conclude that the VCs generated were of a straightforward variety. Though readers interested in a more detailed discussion of how exactly the prover goes about transforming and dispatching VCs are encouraged to refer to \cite{smith:2013}.

@@

Shown below is a RESOLVE facility module that implements the client logic 
\begin{verbatim}
Facility LED_Telos_Demo;
    uses Std_Clock_Fac;

@@

    Operation Main(); Procedure
        Var I1, I2, I3 : Integer;
        Var Loop : Boolean;
        Var L : Led;

@@

Prior to manipulating the LED component developed in the previous sections, we must first pair its specification with a realization. This is accomplished via a facility declaration:

>>>>

In this paper we present a flexible means of specifying and verifying the correctness of software for embedded network systems. Our approach uses RESOLVE, an imperative, component based programming and mathematical specification language, to verify the functional correctness of embedded applications. In doing so, we enrich the work originally presented in \cite{regula:2010} with the following: A model view controller (MVC) based implementation of a RESOLVE to C translator, a dynamic memory allocation scheme tailored towards embedded systems running the code generated by our tool, and the addition of a new language keyword that enables users to pair custom RESOLVE specifications with `externally' implemented (non-native RESOLVE) realizations. We demonstrate these additions on an LED (Light emitting diode) driver that showcases recent mathematical developments, as well as formal verification of a toggling capability enhancement that we demonstrate running on a Telos mote.
\end{abstract}

@@

\subsection{Concepts}
\label{ssec:concepts}
A \textit{concept} module in RESOLVE defines a specification for a mathematical, abstract type. Similar to an interface in Java, a concept provides a number of operation signatures that implementors are expected to realize. Shown below is a \texttt{LED\_Template} concept that provides a light strip abstraction.

@@

It's worth noting that unlike the \texttt{LED\_Template} presented in \cite{regula:2010} which models an LED as the \textit{cartesian product} of booleans $b_1$, $b_2$, \ldots , $b_4$, the strip model we present here instead uses strings for the following reasons:
\begin{itemize}
\item Strings are indexible, and thus do not require separate \texttt{Set} and \texttt{Status} operations for each individual LED.
\item This approach demonstrates the benefits of reusable mathematical theories. The specifications listed here are rooted in \texttt{String\_Theory}, a pre-existing RESOLVE math library that provides most of the definitions and theorems used throughout the specifications.
\end{itemize}

Finally, the concept provides two operations. The first, \texttt{Set}, takes as a parameter an instance of an LED strip \texttt{L}, a boolean \texttt{b}, and an integer \texttt{i}. The operation \texttt{requires} that \texttt{i} falls within the length of the strip, and \texttt{ensures} upon completion that the length of the outgoing strip \texttt{L} is the same as the incoming strip, \texttt{\#L}, and that the LED in position \texttt{i} of \texttt{L} is set to boolean \texttt{b}. The \texttt{Status} operation is specified similarly.

@@

@@

Even enhancement realizations are neutral since only the operations called are those provided by the base concept -- and require no knowledge of how the called operation is implemented.

@@

\begin{figure}[!htb]
\centering
\begin{tabular}{lccc}
	\toprule
	Condition \# & Time (ms)	& Steps	& Search \\
	\midrule
	VC 0\_1	& 4426	& 5	& 0	\\
	VC 0\_2	& 5039	& 5	& 0	\\
	VC 0\_3	& 6324	& 6	& 0	\\
	\bottomrule
\end{tabular}
\caption{Verification results for operation \texttt{Toggle}}
\label{fig:results}
\end{figure}

As the results summarized in Figure \ref{fig:results} indicate, using RESOLVE's integrated prover, we are able to mechanically and automatically dispatch all VCs for \texttt{Toggling\_Realiz}, thus verifying its correctness. In terms of proof difficulty, given the number of steps and time taken to establish each, we may conclude that the VCs generated were of a straightforward variety. Readers interested however in learning more about how the prover goes about transforming and dispatching similar (and other, more complex) VCs should refer to \cite{smith:2013}.

@@

Shown below is a RESOLVE facility module that implements the client logic of our embedded application.
\begin{verbatim}
Facility LED_Telos_Demo;
    uses Std_Clock_Fac;

@@

    Operation Main(); Procedure

        (* Declare LED strip indices *)
        Var I1, I2, I3 : Integer;
        Var Loop : Boolean;

        (* Declare an LED strip *)
        Var L : Led;

@@

Prior to using the LED component developed in the previous sections, we must first pair our \texttt{LED\_Template} specification with an appropriate realization. This is accomplished via the facility declaration located directly beneath the uses clause which pairs our LED specification, \texttt{LED\_Template}, with a ``\texttt{Std\_Led\_Realiz}" realization. In order to give LED strip instances the ability to toggle specific LEDs, we also similarly pair the specification, \texttt{Toggling\_Capability} with our \texttt{Toggling\_Realiz} realization.

The bulk of logic driving the actual application rests in the non terminating busy loop inside operation \texttt{Main}, where we use the operation provided by \texttt{Toggling\_Capability} to turn each light on the strip on then off in succession.

\subsubsection{``External'' Realization Support}
\label{ssec:external}

Readers might note that we never presented a realization of \texttt{LED\_Template}. Indeed, the RESOLVE programmer would ideally, after writing the concept, provide a verifiable native RESOLVE implementation. However, as our target platform is embedded, and our concept aims to provide control for LEDs -- a decidedly low level feature on embedded hardware -- our realization is forced to operate at similarly low levels by manually interfacing with hardware pins provided by msp430 chipset.

RESOLVE, however, in its current state is far too high level of a language to perform these tasks directly -- that is, it currently lacks the driver support to do so. In an effort to address this difficulty, we introduce the notion of \textit{external realizations}, which allows the user to write their own realization of a concept in a language of their choosing. In the case of \texttt{LED\_Template}, we chose to provide a hand written \texttt{Std\_Leds\_Realiz} C realization that matches as closely as possible the conventions and model of translation later detailed in Section \#. An example above shows how users communicate to the compiler that a realization is external.

% Applies to base types in resolve as well

% Similar to NesC, the towers of abstraction idea...

We feel this addition to the language is appropriate and useful for the following reasons:
